Grocery Pricing Strategies: An Analysis of Vendor Positioning and Market Dynamics

### 2024-11-14 paper/paper.qmd
finalize paper

## Sources of Bias
The dataset is subject to several potential biases that may affect the analysis and interpretation of grocery pricing data. First, the data includes only a select number of vendors, possibly excluding smaller, independent stores that may have different pricing and promotional strategies. This limitation could skew our understanding of average pricing, as larger vendors with more data points (e.g., Metro, Walmart) might dominate the analysis, potentially overshadowing strategies that smaller vendors might use. Additionally, the data is region-specific and does not account for regional variations in grocery pricing, meaning that the findings may not fully represent pricing trends across different areas.

Another important source of bias stems from the method of data collection. The data was gathered from online listings, which may differ from in-store pricing and promotions. For instance, in-store promotions or discounts exclusive to loyalty program members would not be captured, potentially underestimating the frequency and diversity of promotions offered by each vendor. This could particularly affect vendors like No Frills and Walmart, which tend to have stable, low prices and may offer in-store promotions not reflected online. Consequently, this method of data collection might result in an incomplete picture of vendor strategies, especially for promotional events and discounts.

>>>>

## Sources of Bias
The dataset is subject to several potential biases that may affect the analysis and interpretation of grocery pricing data. First, the data includes only a select number of vendors, possibly excluding smaller, independent stores that may have different pricing and promotional strategies. This limitation could skew our understanding of average pricing, as larger vendors with more data points (e.g., Metro, Walmart) might dominate the analysis, potentially overshadowing strategies that smaller vendors might use. Additionally, the data is region-specific and does not account for regional variations in grocery pricing, meaning that the findings may not fully represent pricing trends across different areas outside of our target population of vendors within Toronto.

Another important source of bias stems from the method of data collection. The data was gathered from online listings, which may differ from in-store pricing and promotions. For instance, in-store promotions or discounts exclusive to loyalty program members would not be captured, potentially underestimating the frequency and diversity of promotions offered by each vendor. This could particularly affect vendors like No Frills and Walmart, which tend to have stable, low prices and may offer in-store promotions not reflected online. Furthermore, this method of data collection raises the problem of measurement bias, pertaining to in-store pick-up prices. These pick-up prices may differ from those that are available at in-store location or online delivery. The non-random exclusion of such prices bias our estimates for average prices and trends thereof. For instance, consumers choosing in-store pickup often differ in preferences, price sensitivity, and shopping habits from those opting for delivery. Thus, this method may only partially capture vendor strategies, especially regarding promotions and discounts, leading to an incomplete view of pricing dynamics.

The dataset has evolved to capture a broader spectrum of consumer habits by progressively including a wider variety of goods in the basket [@Filipp_2024]. Initially focused on a smaller set of products, the basket later expanded, aiming to better represent the diversity of grocery consumption. However, this inconsistency in product range over time introduces a potential source of bias, particularly in the analysis of price distribution and trends. Including a larger variety of goods can shift the average prices and potentially alter the price distribution across vendors. For instance, if the initial basket emphasized staple goods with relatively stable prices, while later data incorporated specialty items with higher price variability, observed price trends might reflect changes in product range rather than actual market shifts. This inconsistency potentially introduces a bias that renders any perceived trends could stem from the evolution of the dataset itself rather than from underlying price changes. Adjusting for this by creating comparable baskets over time or normalizing based on product categories could help mitigate this bias and provide a clearer understanding of true price patterns.
